---
title: "Project R Markdown"
author: "Evelyn Shi"
date: "`r Sys.Date()`"
output: html_document
---

## Introduction
This data set, known as (Shelby J.) Haberman's Survival Data Set, looks at the 
survival of patients who had undergone surgery for breast cancer. The 
information comes from a study conducted by Billings Hospital
at the University of Chicago from 1958 to 1970. [Haberman](https://www.habermanstatistics.com/) 
is a statistician who was a professor at the University at the time and likely 
compiled this data.

The data can be found [here](https://www.kaggle.com/datasets/gilsousa/habermans-survival-data-set) on Kaggle. 

Included in this data is:

|   1. Patient age at time of operation
|   2. Year of operation (formatted such that 1964 would be `64` in the data set)
|   3. Number of positive axillary nodes* detected
|   4. Survival status where: 
|           1 = the patient survived 5 years or longer
|           2 = the patient died within 5 years
\
*[Axillary nodes](https://www.komen.org/wp-content/uploads/AxillaryLymphNodes-1.pdf) 
are lymph nodes in the underarm. They are the first place breast cancer is 
likely to go if it spreads. During surgery, some of these nodes can be removed 
to see if they contain cancer. 

This report aims to look at the survival of patients after going surgery for 
breast cancer. This will be done by finding the Kaplan Meier Estimate of the 
survival function of the patients.

```{r, loading-data}
# set root directory
here::i_am("report.Rmd")

data <- readRDS(
  file = here::here("output/cleandata.rds")
)
```

## Tabular Analysis
```{r, table}
table <- readRDS(
  file = here::here("output/table.rds")
)
table
```
\
This table is the Kaplan Meier Curve in tabular form. The `Age` column tells us 
the age of the patient at the time of the operation. 
`Number of Patients at Risk` describes the number of patients who had yet to 
undergo the surgery. `Number of Patients Who Died Within 5 Years` is, of 
course, the number of patients who were operated on that year who died 
within 5 years of the operation. `Probability of Survival` caluclates the 
probability of survival after the breast cancer surgery, followed by a 
`Standard Error` and Confidence Interval. 

## Kaplan-Meier Curve
```{r, Kaplan-Meier Curve}
knitr::include_graphics(
  here::here("output/survival.png")
)
```
\
In this plot, the y-axis represents survival probability and the x-axis is 
age at time of breast surgery operation. The dark line is the Kaplan-Meier 
estimate of the survival curve, while the dotted lines are the confidence 
intervals for that estimate. 
